Identification of high-yielding soybean lines with exceptional seed composition qualities

### 2022-06-03 doc/manuscript/manuscript.Rmd
Added more to manuscript. Some changes to add the lsmean plot to the main document

---
title: Identification of high-yielding soybean lines with exceptional seed composition qualities
author:
  - name: Jay Gillenwater
    affil: 1,\ddagger,*
    orcid: 0000-0002-3477-6291
affiliation:
  - num: 1
    address: |
      Department of Crop and Soil Sciences, NCSU, Raleigh, NC, USA
    email: jhgille2@ncsu.edu
# firstnote to eighthnote
firstnote: |
  Current address: Updated affiliation
secondnote: |
  These authors contributed equally to this work.
correspondence: |
  leutnant@fh-muenster.de; Tel.: +XX-000-00-0000.
journal: Agronomy
type: article
status: submit
bibliography: mybibfile.bib
appendix: appendix.tex
simplesummary: |
  A Simple summary goes here.
abstract: |
  In current markets, the primary uses for soybean seed are in products derived from 
  it's oil or protein content. However, growers are compensated based on seed oil so 
  a more valuable crop is one with both high yield, and high protein, oil, or a combination of the two. A negative 
  correlation between seed yield and seed protein makes improving these traits simultaneously difficult but 
  not impossible through conventional breeding. A selection of lines with exceptional yield and seed 
  composition qualities was made from two recombinant inbred line (RIL) soybean mapping populations to identify soybean
  varieties with yield comparable to existing cultivars, and seed composition traits superior to existing cultivars. The performance of these RILs were evaluated in multiple environments and several genotypes were identified with yield comparable to that of existing high-performance cultivars with seed protein and or oil that is equal to or superior to the high-performing cultivars. These genotypes will provide breeders with additional sources of germplasm for continuing efforts to improve seed composition traits, and provide growers with valuable genotype options.
keywords: |
  yield; protein; oil;soybean; protein meal
acknowledgement: |
  All sources of funding of the study should be disclosed. Please clearly 
  indicate grants that you have received in support of your research work. 
  Clearly state if you received funds for covering the costs to publish in open 
  access.
authorcontributions: |
  For research articles with several authors, a short paragraph specifying their 
  individual contributions must be provided. The following statements should be 
  used ``X.X. and Y.Y. conceive and designed the experiments; X.X. performed the 
  experiments; X.X. and Y.Y. analyzed the data; W.W. contributed 
  reagents/materials/analysis tools; Y.Y. wrote the paper.'' Authorship must be
  limited to those who have contributed substantially to the work reported.
conflictsofinterest: |
  The authors declare no conflict of 
  interest.
sampleavailability: |
  Samples of the compounds ...... are available from the authors.
abbreviations:
  - short: MDPI
    long: Multidisciplinary Digital Publishing Institute
  - short: DOAJ
    long: Directory of open access journals
  - short: TLA
    long: Three letter acronym
  - short: LD 
    long: linear dichroism
output: rticles::mdpi_article
---

```{r setup, warning=FALSE,echo = FALSE, message=FALSE}
library(here)
source(here("packages.R"))
tar_load(c(example_table, 
           yield_summary_tables, 
           elite_genotype_tables))
```

@@

@@

These two tests were grown in both the 2020 and 2021 seasons. These tests were grown in CLA and CAS in 2020 and CAS and PLY in 2021. The same phenotypes were evaluated for each genotype in the 2020 and 2021 seasons using the same methodology that was employed in the 2019 season.

@@

Models were fit using the gamem_met function from the metan package[@olivotoMetanPackageMultienvironment2020]. Least-square means (LS Means) for each genotype and trait were calculated using the above model using the emmeans package[@lenthEmmeansEstimatedMarginal2022] in R. The emmeans package was also used to calculate contrasts as a post-hoc test to compare RIL phenotype means to check means for all collected phenotypes. A sidak adjustment was used to account for multiple comparisons in the calculation of contrasts.

Pearson correlation coefficients between each phenotype were calculated with the metan package.

Pearson correlation is calculated for each pair of traits as:

@@

@@

## Genotypes with comparable yield and oil and superior protein
Four genotypes had yield and oil that was similar to that of the yield checks, and seed protein that was greater than the average of the checks. These genotypes are N18-1635 from Test 1 and genotypes N18-1627, N18-1643, and N18-1783 from Test 2.  Summary data for these lines is given in table 1.

@@

```{=tex}
\begin{figure}[H]
\centering
\includegraphics[width=15 cm]{C:/Users/Jay/Documents/R/Projects/Jay_yield/exports/plots/supplemental/supplemental_figure_1.pdf}
\caption{This is a figure, Schemes follow the same formatting. If there are multiple panels, they should be listed as: (\textbf{a}) Description of what is contained in the first panel. (\textbf{b}) Description of what is contained in the second panel. Figures should be placed in the main text near to the first time they are cited. A caption on a single line should be centered.}
\end{figure}
```

>>>>

---
title: Identification of high-yielding soybean lines with exceptional seed composition qualities
author:
  - name: Jay Gillenwater
    affil: 1,\ddagger,*
    orcid: 0000-0002-3477-6291
affiliation:
  - num: 1
    address: |
      Department of Crop and Soil Sciences, NCSU, Raleigh, NC, USA
    email: jhgille2@ncsu.edu
# firstnote to eighthnote
firstnote: |
  Current address: Updated affiliation
secondnote: |
  These authors contributed equally to this work.
journal: Agronomy
type: article
status: submit
bibliography: mybibfile.bib
appendix: appendix.tex
simplesummary: |
  A Simple summary goes here.
abstract: |
  In current markets, the primary uses for soybean seed are in products derived from 
  it's oil or protein content. However, growers are compensated based on seed oil so 
  a more valuable crop is one with both high yield, and high protein, oil, or a combination of the two. A negative 
  correlation between seed yield and seed protein makes improving these traits simultaneously difficult but 
  not impossible through conventional breeding. A selection of lines with exceptional yield and seed 
  composition qualities was made from two recombinant inbred line (RIL) soybean mapping populations to identify soybean
  varieties with yield comparable to existing cultivars, and seed composition traits superior to existing cultivars. The performance of these RILs were evaluated in multiple environments and several genotypes were identified with yield comparable to that of existing high-performance cultivars with seed protein and or oil that is equal to or superior to the high-performing cultivars. These genotypes will provide breeders with additional sources of germplasm for continuing efforts to improve seed composition traits, and provide growers with valuable genotype options.
keywords: |
  yield; protein; oil;soybean; protein meal
acknowledgement: |
  All sources of funding of the study should be disclosed. Please clearly 
  indicate grants that you have received in support of your research work. 
  Clearly state if you received funds for covering the costs to publish in open 
  access.
authorcontributions: |
  For research articles with several authors, a short paragraph specifying their 
  individual contributions must be provided. The following statements should be 
  used ``X.X. and Y.Y. conceive and designed the experiments; X.X. performed the 
  experiments; X.X. and Y.Y. analyzed the data; W.W. contributed 
  reagents/materials/analysis tools; Y.Y. wrote the paper.'' Authorship must be
  limited to those who have contributed substantially to the work reported.
conflictsofinterest: |
  The authors declare no conflict of 
  interest.
sampleavailability: |
  Samples of the compounds ...... are available from the authors.
abbreviations:
  - short: RIL
    long: Recombinant inbred line
  - short: RCBD
    long: Randomized complete block design
  - short: CLA
    long: Central crops research station
  - short: CAS
    long: Caswell research farm
  - short: PLY  
    long: Tidewater research station
output: rticles::mdpi_article
---

```{r setup, warning=FALSE,echo = FALSE, message=FALSE}
library(here)
source(here("packages.R"))
tar_load(c(example_table, 
           yield_summary_tables, 
           elite_genotype_tables, 
           phenotype_summary_plots))
```

@@

## Experimental design

@@

These two tests were grown in both the 2020 and 2021 seasons. These tests were grown in CLA and CAS in 2020 and CAS and PLY in 2021. RILs were grown in a randomized complete block design with four replications in each location. The same phenotypes were evaluated for each genotype in the 2020 and 2021 seasons using the same methodology that was employed in the 2019 season.

@@

Models were fit using the gamem_met function from the metan package[@olivotoMetanPackageMultienvironment2020]. Least-square means (LS Means) for each genotype and trait were calculated using the above model using the emmeans package[@lenthEmmeansEstimatedMarginal2022] in R. The emmeans package was also used to calculate contrasts as a post-hoc test to compare RIL phenotype means to check means for all collected phenotypes. A sidak adjustment was used to account for multiple comparisons in the calculation of contrasts.Pearson correlation coefficients between each phenotype were calculated with the metan package. Pearson correlation coefficients between each phenotype were calculated with the metan package as well.

The pearson correlation is calculated for each pair of traits as:

@@

```{r, lsmean_plot_link, echo = FALSE}

# Get the link to the lsmean scatterplot
mean_plot_link <- phenotype_summary_plots[grepl("lsmean_scatterplot", phenotype_summary_plots)]

```{=tex}
\begin{figure}[H]
\centering
\includegraphics[width=15 cm]{`r mean_plot_link`}
\caption{Genotype Least square means for seed oil, seed protein, and seed yield for soybean RILs in Test 1 (\textbf{A}) and Test 2 (\textbf{B}). Points indicate the least square means and error bars indicate the standard errors in the estimation of least square means. Blue dots indicate that a RIL has a least square mean above the average of the checks in the test and a color of red indicates that the RIL has a mean value below the check average. The check average is shown with the vertical dashed line.}
\end{figure}
```

@@

## Genotypes with comparable seed yield and seed oil and superior seed protein content
Four genotypes had yield and oil that was similar to that of the yield checks, and seed protein that was greater than the average of the checks. These genotypes are N18-1635 from Test 1 and genotypes N18-1627, N18-1643, and N18-1783 from Test 2.  Summary data for these lines is given in table 1.

@@

### 2022-06-03 doc/manuscript/manuscript.tex
Added more to manuscript. Some changes to add the lsmean plot to the main document

% Affiliations / Addresses (Add [1] after \address if there is only one affiliation.)
\address{%
$^{1}$ \quad Department of Crop and Soil Sciences, NCSU, Raleigh, NC,
USA; \href{mailto:jhgille2@ncsu.edu}{\nolinkurl{jhgille2@ncsu.edu}}\\
}
% Contact information of the corresponding author
\corres{Correspondence: \href{mailto:leutnant@fh-muenster.de}{\nolinkurl{leutnant@fh-muenster.de}};
Tel.: +XX-000-00-0000.}

@@

@@

Models were fit using the gamem\_met function from the metan
package\citep{olivotoMetanPackageMultienvironment2020}. Least-square
means (LS Means) for each genotype and trait were calculated using the
above model using the emmeans
package\citep{lenthEmmeansEstimatedMarginal2022} in R. The emmeans
package was also used to calculate contrasts as a post-hoc test to
compare RIL phenotype means to check means for all collected phenotypes.
A sidak adjustment was used to account for multiple comparisons in the
calculation of contrasts.

Pearson correlation coefficients between each phenotype were calculated
with the metan package.

Pearson correlation is calculated for each pair of traits as:

@@

@@

\hypertarget{genotypes-with-comparable-yield-and-oil-and-superior-protein}{%
\subsection{Genotypes with comparable yield and oil and superior
protein}\label{genotypes-with-comparable-yield-and-oil-and-superior-protein}}

@@

\begin{figure}[H]
\centering
\includegraphics[width=15 cm]{C:/Users/Jay/Documents/R/Projects/Jay_yield/exports/plots/supplemental/supplemental_figure_1.pdf}
\caption{This is a figure, Schemes follow the same formatting. If there are multiple panels, they should be listed as: (\textbf{a}) Description of what is contained in the first panel. (\textbf{b}) Description of what is contained in the second panel. Figures should be placed in the main text near to the first time they are cited. A caption on a single line should be centered.}
\end{figure}

@@

\noindent
\begin{tabular}{@{}ll}
MDPI & Multidisciplinary Digital Publishing Institute \\
DOAJ & Directory of open access journals \\
TLA & Three letter acronym \\
LD & linear dichroism \\
\end{tabular}}

>>>>

% Affiliations / Addresses (Add [1] after \address if there is only one affiliation.)
\address{%
$^{1}$ \quad Department of Crop and Soil Sciences, NCSU, Raleigh, NC,
USA; \href{mailto:jhgille2@ncsu.edu}{\nolinkurl{jhgille2@ncsu.edu}}\\
}
% Contact information of the corresponding author
\corres{Correspondence: }

@@

\hypertarget{experimental-design}{%
\subsection{Experimental design}\label{experimental-design}}

@@

Models were fit using the gamem\_met function from the metan
package\citep{olivotoMetanPackageMultienvironment2020}. Least-square
means (LS Means) for each genotype and trait were calculated using the
above model using the emmeans
package\citep{lenthEmmeansEstimatedMarginal2022} in R. The emmeans
package was also used to calculate contrasts as a post-hoc test to
compare RIL phenotype means to check means for all collected phenotypes.
A sidak adjustment was used to account for multiple comparisons in the
calculation of contrasts.Pearson correlation coefficients between each
phenotype were calculated with the metan package. Pearson correlation
coefficients between each phenotype were calculated with the metan
package as well.

The pearson correlation is calculated for each pair of traits as:

@@

\begin{figure}[H]
\centering
\includegraphics[width=15 cm]{C:/Users/Jay/Documents/R/Projects/Jay_yield/exports/plots/lsmean_scatterplot.pdf}
\caption{Genotype Least square means for seed oil, seed protein, and seed yield for soybean RILs in Test 1 (\textbf{A}) and Test 2 (\textbf{B}). Points indicate the least square means and error bars indicate the standard errors in the estimation of least square means. Blue dots indicate that a RIL has a least square mean above the average of the checks in the test and a color of red indicates that the RIL has a mean value below the check average. The check average is shown with the vertical dashed line.}
\end{figure}

@@

\hypertarget{genotypes-with-comparable-seed-yield-and-seed-oil-and-superior-seed-protein-content}{%
\subsection{Genotypes with comparable seed yield and seed oil and
superior seed protein
content}\label{genotypes-with-comparable-seed-yield-and-seed-oil-and-superior-seed-protein-content}}

@@

@@

\noindent
\begin{tabular}{@{}ll}
RIL & Recombinant inbred line \\
RCBD & Randomized complete block design \\
CLA & Central crops research station \\
CAS & Caswell research farm \\
PLY & Tidewater research station \\
\end{tabular}}
